---
title: "RoC_Maps"
author: "Walter Finsinger"
date: "10/17/2016"
output: html_document
---

This script makes gridded maps of rate-of-vegetation change (RoC) at time lags +500 years from baseline age for selected time slices.
It uses output from the "EPD_Novelty_RoC.Rmd" script that is stored in the directory with the path:
./output.dir/Novelty_Maps

**NB: before running the script, change the "dir" on Line 23 accordingly.**

Loads packages
```{r}
require(paleofire)
require(gridExtra)
require(ggplot2)
```

Create output directory where data and figures are stored
```{r}
dir = "./RoCchord_Novchord_ListFALSE_Start15000_kth1/"
#output.dir = file.path(".",dir)

map.output.dir = paste(dir,"RoC_Maps/", sep='')
dir.create(map.output.dir)
```

Loads RoC data from the DataOut directory
```{r}
rofc_geo <- read.csv(file=paste(dir,"DataOut/rofc_geo.csv", sep=''), header=T, row.names=NULL)
rofc_geo <- rofc_geo[ ,-1] # omits first column, which is equivalent to the row.numbers
```

Modifies the rofc_geo data.frame
```{r}
n.row = nrow(rofc_geo)
n.col = length(rofc_geo)-4
cds = matrix(NA, nrow=1, ncol=6)
colnames(cds) = c("ent", "lon", "lat", "alt", "CD", "agebin")
cds = as.data.frame(cds)
myTS = seq(500,15000,by=500)

for (i in 1:n.col) {
  cdsi = rofc_geo[ ,c(1:4, i+4)]
  colnames(cdsi) [5] = "CD"
  cdsi$agebin = myTS[i]
  cds = rbind(cds, cdsi)
}

cds.map <- cds[which(complete.cases(cds)), ]
```

## Modifies cds data.frame such that it can be used by paleofire mapping tool
```{r}
cds.map = cds.map[ ,-c(1,4)]
cds.map = cds.map[ ,c(1,2,4,3)]
colnames(cds.map) = c("x", "y", "age", "char")
```

## Define TS.Ages for stacked images
```{r}
TS.Age = seq(from=500, to=15000, by=500)
n.TS.Age = length(TS.Age)
end.Bin.out <- max(cds.map$age) + 500
```

Sets parameters used to produce the gridded maps
```{r}
title.size = 10
map.pal <- "YlGn"
col.class = quantile(cds.map$char)
dist.buffer = 300000
cell.size = 50000
x.maplim=c(-800000, 2800000)
y.maplim=c(3700000, 7500000)
```


Maps of RoC 500 yrs lag: Plots Quartile maps on one page for AgeBins younger than that defined by the user-defined parameter 'end.Bin.out', which determines the youngest baseline age that had been excluded from the analysis.
```{r}
Map=TS.Age[1] # TS=500 cal BP
p.map = pfGridding(cds.map, age=Map,
                   distance_buffer=dist.buffer, elevation_buffer=500)
pp.i = plot.pfGridding(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.1 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.1)

pp.legend = pp.i +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="right")

Map=TS.Age[2] # TS=1000 cal BP
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.2 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.2)

Map=TS.Age[5] # TS = 2500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.3 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")

Map=TS.Age[9] # TS=4500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.4 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")

Map=TS.Age[13] # TS=6500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.5 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")

Map=TS.Age[17] # TS = 8500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4, pp.5)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.6 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")

Map=TS.Age[21] # TS=10500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.7 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")

Map=TS.Age[23] # TS = 11500 cal BP
  if(end.Bin.out - 500 < Map) {
    pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
    grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, ncol=3, nrow=4)
    dev.off()
  
    pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
    plot(pp.legend)
    dev.off()
    rm(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7)
    stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
  } else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.8 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.8)

Map=TS.Age[25] # TS = 12500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.9 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.9)

Map=TS.Age[27] # TS = 13500 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.10 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.10)

Map=TS.Age[28] # TS = 14000 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, pp.10, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, pp.10)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.11 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.11)

Map=TS.Age[29] # TS = 14000 cal BP
if(end.Bin.out - 500 < Map) {
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
  grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, pp.10, pp.11, ncol=3, nrow=4)
  dev.off()
  
  pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
  plot(pp.legend)
  dev.off()
  rm(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, pp.10, pp.11)
  stop(paste("Older maps cannot be plotted: oldest baseline age > next Map age (",Map,"cal BP)"), sep=NULL)
} else { }
p.map = pfGridding(cds.map, age=Map, distance_buffer=300000, elevation_buffer=500)
pp.i = plot(x=p.map,
            continuous=F,
            col_class=col.class,
            points=F,
            xlim=x.maplim, ylim=y.maplim,
            cpal=map.pal)
pp.12 = pp.i + ggtitle(paste("RoC", Map,"->",Map-500,"cal BP", sep=" ")) +
  geom_point(data=p.map$points, aes(x=x,y=y), size = 0.5, colour="black") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  theme(plot.title = element_text(size=title.size, hjust=0.5)) +
  theme(legend.position="none")
plot(pp.12)


pdf(file.path(map.output.dir, "RoC_Maps_quartiles.pdf"))
grid.arrange(pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, pp.10, pp.11, pp.12, ncol=3, nrow=4)
dev.off()

pdf(file.path(map.output.dir, "RoC_Maps_quartiles_Legend.pdf"))
plot(pp.legend)
dev.off()

rm(pp.i, p.map, pp.1, pp.2, pp.3, pp.4, pp.5, pp.6, pp.7, pp.8, pp.9, pp.10, pp.11, pp.12, pp.legend)
```
