---
title: "EPD RoC Models Pre 8K"
author: "Simon Brewer"
date: "October 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data 
```{r}
require(reshape2)
require(nlme)
require(mgcv)

input.dir = paste("../RoCchord_Novchord_ListFALSE_Start15000_kth1/DataOut/", sep='_')

## Loads data ####

## NGRIP data
ngrip <- read.table(file="http://www.iceandclimate.nbi.ku.dk/data/2010-11-19_GICC05modelext_for_NGRIP.txt", skip=70)
#ngrip <- read.csv(file="./Other_Data/2010-11-19_GICC05modelext_for_NGRIP.csv", header=T)
colnames(ngrip) <- c("Age_b2k", "depth", "d18O", "MCE")
# corrects ngrip$Age_b2k...NGRIP uses b2k age scale, where 50 b2k = 0 cal yr BP (AD 1950)
ngrip$Age <- ngrip$Age_b2k - 50   

# Plot checks age boundaries of YD (as from Table 2 in Rasmussen et al. (2014):
# Rasmussen, S.O., Bigler, M., Blockley, S.P., Blunier, T., Buchardt, S.L.,
# Clausen, H.B., et al. (2014). A stratigraphic framework for abrupt climatic changes
# during the Last Glacial period based on three synchronized Greenland ice-core
# records: refining and extending the INTIMATE event stratigraphy. Quaternary Science Reviews, 106, 14–28
# Start of GS-1: 12,896 ± 4 years b2k => 12,846 cal BP
# Start of Holocene: 11,703 ± 4 years b2k => 11,653 cal BP
# Start of GI-1 (Bolling): 14,692 ± 4 years B2k => 14,642 cal BP
par(mfrow=c(1,1))
plot(ngrip$Age, ngrip$d18O, xlim=rev(range(c(15000, 0))), type="l")
abline(v = c(14642, 12846, 11653))

## Bin up the the NGRIP data
time.mid = seq(0, 15500,500)
d18O.bin = rep(NA, length(time.mid))
d18O.bin.sd = rep(NA, length(time.mid))
half.win = 250
for (i in 1:length(time.mid)) {
  age.lo = time.mid[i] - half.win
  age.hi = time.mid[i] + half.win
  ageID = which(ngrip$Age >= age.lo & ngrip$Age < age.hi)
  d18O.bin[i] = mean(ngrip$d18O[ageID], na.rm=TRUE)
  d18O.bin.sd[i] = sd(ngrip$d18O[ageID], na.rm=TRUE)
}
d18O.bin <- as.data.frame(d18O.bin)
colnames(d18O.bin) <- "d18O"
d18O.bin$Age <- time.mid
d18O.bin$d18O.diff = c(NA, abs(diff(d18O.bin$d18O))) # gives differenced d18O for baseline Age -> subsequent target Age
                                                # Age in d18O.bin = TS.Age in other records!!!
d18O.bin$d18O.diff = log(d18O.bin$d18O.diff)
d18O.bin$d18O.sd = d18O.bin.sd
d18O.bin <- d18O.bin[-c(1,32), ]

lines(d18O.bin$Age,d18O.bin$d18O, type='l', lwd=3, col=2)

## differenced binned NGRIP 
# d18O.bin <- read.csv(paste(input.dir,"d18O_bin.csv", sep=''))
# d18O.bin <- d18O.bin[-c(1,32), ]

## ---- DATA ---- ##
## Rate of change record
rofc_geo <- read.csv(paste(input.dir,"rofc_geo.csv", sep=''), header=T, row.names = NULL)
rofc_geo <- rofc_geo[, -1]

## Median novelty at 500 year lags
Nov500 <- read.csv(paste(input.dir,"Nov500.csv", sep=''))

## All novelty values at 500 year lags
Nov500lag <- read.csv(paste(input.dir,"Nov500lag.csv", sep=''))

## Check data with plots
par(mfrow=c(2,1), mar=c(1,5,0.5,1), oma=c(5,1,1,1), cex=0.7)
plot(d18O.bin$Age, d18O.bin$d18O.diff, type="o", xaxt="n", ylab="d18O change")
abline(v = c(14642, 12846, 11653))
plot(Nov500$TS.Age, Nov500$med.cma.dist, type="o", ylab="median novelty")
abline(v = c(14642, 12846, 11653))

par(mfrow=c(1,2))
plot(d18O.bin$d18O.diff[13:30], Nov500$med.cma.dist[13:30], 
     xlab="d18O change", ylab="Med novelty")
plot(d18O.bin$d18O.sd[13:30], Nov500$med.cma.dist[13:30], 
     xlab="d18O change", ylab="Med novelty")
```

## Gets the RoC data into shape

```{r}
time.mid <- seq(500, 15000, 500)
## Melt RoC data.frame. NB: agepoll = age of baseline set
cds = melt(rofc_geo, id.vars = c("ent","lon","lat","alt"),
           variable.name = "agepoll", 
           value.name = "cd")
## Modifies the cds data.frame (writes baselines' ages, and binned NGRIP d18O values)
cds$ent = as.factor(cds$ent)
cds$agebp = rep(time.mid, each=dim(rofc_geo)[1]) # agebp = age of baseline set
cds$d18O = rep(d18O.bin$d18O.diff, each=dim(rofc_geo)[1]) # wf added a minus sign in d18O.bin
#cds$d18O <- abs(cds$d18O)
cds$d18O.sd <- rep(d18O.bin$d18O.sd, each=dim(rofc_geo)[1])

cds = cds[which(complete.cases(cds)),]

## Log transform CD scores
cds$lcd = log(cds$cd)

par(mfrow=c(1,1))
plot(cds$agebp, cds$cd)
plot(cds$agebp, cds$lcd, type="o") # Note there's an outlier at 2000 cal BP having an lcd < -6!!
outl <- cds[which(cds$lcd < -6), ]

## Delete the one outlier sample
cds <- cds[-which(cds$lcd < -6), ]
plot(cds$agebp, cds$lcd, type="o")
boxplot(cds$lcd ~ cds$agebp)

## Pre 8K only
cds.lg = subset(cds, agebp>8000)

## mean center
cds.lg$agebp.c = cds.lg$agebp - mean(cds.lg$agebp)
cds.lg$d18O.c = cds.lg$d18O - mean(cds.lg$d18O)
```

# Models

## OLS model
```{r}
scd.ols = lm(lcd ~ 1, data=cds.lg)
summary(scd.ols)
```

## Unconditional means model (LME)
```{r}
scd.fit0 = lme(lcd ~ 1, random=~1|ent, data=cds.lg, method="ML")
summary(scd.fit0)
```

Maximum likelihood fit model with AIC of `r round(AIC(scd.fit0),2)`.

Fixed effects (significant under Wald test):
```{r}
fixed.effects(scd.fit0)
```

Variance components
```{r}
VarCorr(scd.fit0)
```

- Variance of intercept/between site variability ($\tau^2$): `r round(as.numeric(VarCorr(scd.fit0)[1,1]),4)`
- Variance within cores/residuals ($\sigma^2$): `r round(as.numeric(VarCorr(scd.fit0)[2,1]),4)`
- Approx `r as.numeric(VarCorr(scd.fit0)[1,1])/(as.numeric(VarCorr(scd.fit0)[1,1])+as.numeric(VarCorr(scd.fit0)[2,1]))*100` percent between sites

## Random intercepts model (LME)
```{r}
scd.fit1 = lme(lcd ~ agebp, random=~1|ent, data=cds.lg, method="ML")
summary(scd.fit1)
```

Maximum likelihood fit model with AIC of `r round(AIC(scd.fit1),2)`.

Fixed effects (significant under Wald test):
```{r}
fixed.effects(scd.fit1)
```

Variance components
```{r}
VarCorr(scd.fit1)
```

- Variance of intercept/between site variability ($\tau^2$): `r round(as.numeric(VarCorr(scd.fit1)[1,1]),4)`
- Variance within cores/residuals ($\sigma^2$): `r round(as.numeric(VarCorr(scd.fit1)[2,1]),4)`
- Approx `r as.numeric(VarCorr(scd.fit1)[1,1])/(as.numeric(VarCorr(scd.fit1)[1,1])+as.numeric(VarCorr(scd.fit1)[2,1]))*100` percent between sites

Now examine AIC to see whether including the random intercepts has helped (yes!):
```{r}
sapply(list(scd.ols,scd.fit0,scd.fit1), AIC)
```

Pseudo $r^2$ ((sigma of unconditional means) - (sigma of random intercept)) / (sigma of unconditional means)
```{r}
(as.numeric(VarCorr(scd.fit0)[2,1]) - as.numeric(VarCorr(scd.fit1)[2,1]))/ as.numeric(VarCorr(scd.fit0)[2,1])
```
About 3% of variance of LCD by relationship with AGEBP. Therefore single linear relationship (i.e. fixed slope) is not a good model

## Random slopes and intercepts model (LME)
```{r}
scd.fit2 = lme(lcd ~ agebp.c, random=~1+agebp.c|ent, data=cds.lg, method="ML")
summary(scd.fit2)
```

Maximum likelihood fit model with AIC of `r round(AIC(scd.fit2),2)`.

Fixed effects (significant under Wald test):
```{r}
fixed.effects(scd.fit2)
```

Variance components
```{r}
VarCorr(scd.fit2)
```

- Variance of intercept/between site variability ($\tau^2_0$): `r round(as.numeric(VarCorr(scd.fit2)[1,1]),4)`
- Variance in site response over time ($\tau^2_1$): `r round(as.numeric(VarCorr(scd.fit2)[2,1]),4)`
- Variance within cores/residuals ($\sigma^2$): `r round(as.numeric(VarCorr(scd.fit2)[3,1]),4)`
- Approx `r as.numeric(VarCorr(scd.fit2)[1,1])/(as.numeric(VarCorr(scd.fit2)[1,1])+as.numeric(VarCorr(scd.fit2)[3,1]))*100` percent between sites

Now examine AIC to see whether including the random intercepts has helped (yes!):
```{r}
sapply(list(scd.ols,scd.fit0,scd.fit1,scd.fit2), AIC)
```

## Random [slopes and intercepts] model w/ d18O (LME)
```{r}
scd.fit3 = lme(lcd ~ d18O.c, random=~1|ent, data=cds.lg, method="ML")
summary(scd.fit3)
```

Maximum likelihood fit model with AIC of `r round(AIC(scd.fit3),2)`.

Fixed effects (significant under Wald test):
```{r}
fixed.effects(scd.fit3)
```

Variance components
```{r}
VarCorr(scd.fit3)
```

- Variance of intercept/between site variability ($\tau^2_0$): `r round(as.numeric(VarCorr(scd.fit3)[1,1]),4)`
- Variance in site response to d18O ($\tau^2_1$): `r round(as.numeric(VarCorr(scd.fit3)[2,1]),4)`
- Approx `r as.numeric(VarCorr(scd.fit3)[1,1])/(as.numeric(VarCorr(scd.fit3)[1,1])+as.numeric(VarCorr(scd.fit3)[2,1]))*100` percent between sites

Now examine AIC to see whether including the random intercepts has helped (yes!):
```{r}
sapply(list(scd.ols,scd.fit0,scd.fit1,scd.fit2,scd.fit3), AIC)
```

## Additive random slopes and intercepts model w/ d18O (LME)
```{r}
scd.fit4 = gamm(lcd ~ agebp.c + s(d18O, k=4), random=list(ent=~1+d18O), data=cds.lg, method="ML")
summary(scd.fit4$lme)
```

Maximum likelihood fit model with AIC of `r round(AIC(scd.fit3),2)`.

Fixed effects (significant under Wald test):
```{r}
fixed.effects(scd.fit4$lme)
```

Variance components
```{r}
VarCorr(scd.fit4$lme)
```

- Variance of intercept/between site variability ($\tau^2_0$): `r round(as.numeric(VarCorr(scd.fit3)[1,1]),4)`
- Variance in site response to d18O ($\tau^2_1$): `r round(as.numeric(VarCorr(scd.fit3)[2,1]),4)`
- Approx `r as.numeric(VarCorr(scd.fit3)[1,1])/(as.numeric(VarCorr(scd.fit3)[1,1])+as.numeric(VarCorr(scd.fit3)[2,1]))*100` percent between sites

Now examine AIC to see whether including the random intercepts has helped (yes!):
```{r}
sapply(list(scd.ols,scd.fit0,scd.fit1,scd.fit2,scd.fit3), AIC)

save(cds.lg,scd.ols,scd.fit0,scd.fit1,scd.fit2,scd.fit3, file="RoC_Pre8K.RData")
```


